Based on 2022 paper, "" Does culture moderate relationships between rumination and symptoms of depression and PTSD? 

-overall rumination rates
 -difference by culture
                income
                age
                religion 
                gender
                education
                *trauma type (for PTSD)


```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(paletteer)

```

```{r}
dep<-read_csv(here("jka_r_project", "rum_depress_jka.csv")) %>% clean_names
```
```{r}
my_palette <- paletteer_d("ggprism::stained_glass2")
```


```{r}
ptsd<-read_csv(here("jka_r_project", "rum_ptsd_jka.csv")) %>% clean_names
```

```{r}
glimpse (dep)
```
```{r}
na.omit(dep)
```
```{r}
select()
```


Below is how I identified culture = 0 as Australian, based on the averge ages of the two groups in the first table of the paper. 

```{r}
dep %>% 
  filter(culture==0) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```
```{r}
dep %>% 
  select(id, xpcult, gender, age)
```

This shows that the first 109 study participants were Australian, with the remainder being Malaysian. At id 109, culture changes from 0 to 1.  

```{r}

dep$xpcult <- ifelse(dep$culture == 0, "Euro-Australian", "Malaysian")
dep$xpcult
```

```{r}
dep$bingen <- ifelse(dep$gender == 1, "M", "F")
dep$bingen
```


```{r}
glimpse(dep)
```
```{r}
dep %>% 
  na.omit() %>% 
  select(age, bingen, education, xpcult, religion_coded, depression_total, anxiety_total, brooding_total)
```



```{r}
dep %>% 
  filter(between(religion_coded, 0, 6))%>% 
         group_by(xpcult) %>% 
  summarize(avg_brood = mean(brooding_total)) %>% 
  ggplot(aes(x = xpcult, y = avg_brood, fill = xpcult))+
  geom_col(alpha=0.5)+
  scale_fill_manual(values=my_palette)+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size=10)) +
  labs(title = "Brooding by Culture",
       x = "Culture",
       y = "Average Brooding",
       fill = "Culture")
```


```{r}
dep %>% 
  mutate(education = as.factor(education)) %>% 
  group_by(education) %>% 
  summarize(avg_brood = as.factor(mean(brooding_total))) %>% 
  ggplot(aes(x = education, y = avg_brood, fill = education))+
  geom_col(alpha=0.5)+
  scale_fill_manual(values=my_palette)+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size=10)) +
  labs(title = "Brooding by Education",
       x = "Education",
       y = "Average Brooding")

```

```{r}


```


```{r}
glimpse(ptsd)
```




```{r}
glimpse(ptsd)
```

```{r}
names(ptsd)
```


