Based on 2022 paper, "https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0278328" Does culture moderate relationships between rumination and symptoms of depression and PTSD? 

-overall rumination rates
 -difference by age... to overal rumination rates (brooding total)
 
 Facet by culture(only 2 diff cultures mala and euro australian)
              
                income
                age
              *trauma type (for PTSD)
                
                Mine to do:
                religion 
                gender
                education(on a 1-5 scale )
                
Gender:              
 M=1
 F=2
Declined to state  =5  

Culture:  
Australian = 0
Malaysian = 1

Education (Highest level achieved):
Secondary = 1
Post-secondary = 2
Undergrad degree = 3
Graduate degree = 4
Other = 5

Religion code (for depression dataset):
None = 1
Christian = 2
Muslim = 3
Buddhist/Taoist = 4
Hindu = 5
Other (incl Sikh, Agnostic) = 6

Index trauma type (what the PTSD is thought to be in response to; for PTSD dataset):
Accident/serious injury/illness = 1
Non-sexual assault/abuse = 2
Sexual assault/abuse = 3
Witnessing death = 4
War/natural disaster = 5
Other = 6                



ptsd data >> DO NOT LOOK AT "HADS"
>we want to look at pts (pcl 1 is q 1) total and if that value is above a certian value, then they have ptsd>> mutate boolean t/f 
>rrs is the brooding score sum the total and *maybe* append to depress to see total brooding score with depress  

So we are looking at rumination rates in the Y axis, and X axis is relig, gender, education...
and we have a facet of culture
```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)

```

```{r}
depress<-read_csv(here("R_project", "Rumination Depression.csv")) %>% clean_names
```


```{r}
ptsd<-read_csv(here("R_project","Rumination PTSD.csv")) %>% clean_names
```

```{r}
glimpse (ptsd)
```

```{r}
naniar::miss_var_summary(depress)
```


```{r}
glimpse(ptsd)
```


```{r}
naniar::miss_var_summary(ptsd)
```
```{r}
depress %>% 
  summary(culture)
  ggplot(aes(x=age, y=mean(brooding_total, na.rm=T)))+geom_col(na.rm = T)+facet_wrap(culture~.)

```


