Based on 2022 paper, 'https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0278328' Does culture moderate relationships between rumination and symptoms of depression and PTSD? 

-overall rumination rates
 -difference by age... to overal rumination rates (brooding total)
 
 Facet by culture(only 2 diff cultures mala and euro australian)
              
                income
                age
              *trauma type (for PTSD)
                
                Mine to do:
                religion 
                gender
                education(on a 1-5 scale )
                
Gender:              
 M=1
 F=2
Declined to state  =5  

Culture:  
Australian = 0
Malaysian = 1

Education (Highest level achieved):
Secondary = 1
Post-secondary = 2
Undergrad degree = 3
Graduate degree = 4
Other = 5

Religion code (for depression dataset):
None = 1
Christian = 2
Muslim = 3
Buddhist/Taoist = 4
Hindu = 5
Other (incl Sikh, Agnostic) = 6

Index trauma type (what the PTSD is thought to be in response to; for PTSD dataset):
Accident/serious injury/illness = 1
Non-sexual assault/abuse = 2
Sexual assault/abuse = 3
Witnessing death = 4
War/natural disaster = 5
Other = 6                



ptsd data >> DO NOT LOOK AT "HADS"
>we want to look at pts (pcl 1 is q 1) total and if that value is above a certian value, then they have ptsd>> mutate boolean t/f 
>rrs is the brooding score sum the total and *maybe* append to depress to see total brooding score with depress  

So we are looking at rumination rates in the Y axis, and X axis is relig, gender, education...
and we have a facet of culture
```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)

```

```{r}
depress<-read_csv(here("R_project", "Rumination Depression.csv")) %>% clean_names
```


```{r}
post_disorder<-read_csv(here("R_project","Rumination PTSD.csv")) %>% clean_names
```

```{r}
glimpse(ptsd)
```

```{r}
naniar::miss_var_summary(depress)
```


```{r}
glimpse(post_disorder)
```


```{r}
naniar::miss_var_summary(ptsd)
```
Making the dataset readable

```{r}

#write.csv( d, "Depression_data", row.names=FALSE)

d<-depress %>% 
  select(-age,-brooding_total,-anxiety_total,-depression_total) %>% 
  mutate_if(is.numeric,as.character )
d<-d %>% mutate(age=depress$age)
d<-d %>% mutate(brooding_total=depress$brooding_total)
d<-d %>% mutate(anxiety_total=depress$anxiety_total)
d<-d %>% mutate(depression_total=depress$depression_total)
new_education<-d %>% select(education) %>% 
mutate_if( is.character, as.factor)

```
  To their names
```{r}

  
d["culture"][d["culture"]==0]<-"Australian"
d["culture"][d["culture"]==1]<-"Malaysian"  
d["gender"][d["gender"]==1]<-"M"
d["gender"][d["gender"]==2]<-"F"
d["gender"][d["gender"]==5]<-"Declined to State"
d["education"][d["education"]==1]<-"Secondary"
d["education"][d["education"]==2]<-"Post-secondary"
d["education"][d["education"]==3]<-"Undergrad degree"
d["education"][d["education"]==4]<-"Graduate degree"
d["education"][d["education"]==5]<-"Other"
d["religion_coded"][d["religion_coded"]==1]<-"None"
d["religion_coded"][d["religion_coded"]==2]<-"Christian"
d["religion_coded"][d["religion_coded"]==3]<-"Muslim"
d["religion_coded"][d["religion_coded"]==4]<-"Buddhist/Taoist"
d["religion_coded"][d["religion_coded"]==5]<-"Hindu"
d["religion_coded"][d["religion_coded"]==6]<-"Other (incl Sikh, Agnostic)"
d<-d %>% 
  select(-education) %>% mutate(education=new_education$education)
d
```


















Debugging:


Based on 2022 paper, "https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0278328" Does culture moderate relationships between rumination and symptoms of depression and PTSD? 
Brooding/RRS rates & culture with: 

Akshanth:      religion, gender, education(1-5; change to factors) 

JKA:           depression, anxiety, age, trauma type for PTSD


----------------------------------------------------------------------------------------------

Data codes for variables of interest from paper:

For both: First 109 participant ids are Euro-Australian, next 144 are Malaysian

For depression data:

2 Cultures:  0,1 = Euro-Australian, Malaysian (convert to factors)

Gender (convert to factors):              
 M = 1
 F = 2
Declined to state = 5  

Education (Highest level achieved):
Secondary = 1
Post-secondary = 2
Undergrad degree = 3
Graduate degree = 4
Other = 5

Religion codes:
None = 1
Christian = 2
Muslim = 3
Buddhist/Taoist = 4
Hindu = 5
Other (incl Sikh, Agnostic) = 6

PTSD dataset:

Culture = Groups 1,2  = Euro-Australian, Malaysian (change to factors)

Index trauma type (what the PTSD is thought to be in response to):

Accident/serious injury/illness = 1
Non-sexual assault/abuse = 2
Sexual assault/abuse = 3
Witnessing death = 4
War/natural disaster = 5
Other = 6       

HADS = Hospital Anxiety and Depression Scale: 

had_sdep = depression subscale total
had_sanx = anxiety subscale total
had_stotal = overall HADS total

pts_dtotal = overall score on PTSD assessment scale

RRS-B = Brooding scale - want sum of all RRS scores

```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(paletteer)
```
```{r}
my_palette <- paletteer::paletteer_d(("ggprism::colors"))
```

```{r}
dep<-read_csv("jka_r_project/rum_depress_jka.csv") %>% clean_names
```

```{r}
ptsd<-read_csv("jka_r_project/rum_ptsd_jka.csv") %>% clean_names
```

```{r}
glimpse(dep)
```

````{r}
dep$culture <- as.factor(dep$culture)
class(dep$culture)
```


```{r}
dep %>% 
  filter(culture==0) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```

```{r}
dep %>% 
  filter(culture==1) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```

```{r}
dep <- dep %>% 
  mutate(age_cat = case_when(age < 21 ~ "teens",
                             age >= 21 & age < 25 ~ "earlytwens",
                             age >= 25 & age < 30 ~ "latetwens",
                             age >= 30 & age < 35 ~ "earlythirs",
                             age >= 35 & age < 40 ~ "latethirs",
                             age >= 40 & age < 45 ~ "earlyfors",
                             age >= 45 & age < 50 ~ "latefors",
                             age >= 50 & age < 55 ~ "earlyfifs",
                             age >= 55 & age < 60 ~ "latefifs",
                             age >= 60 ~ "sixties"))
```

```{r}
dep %>%
  na.omit() %>%
  group_by(culture, age_cat) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T)) %>%
  select(culture, age_cat, avg_brood)
```




```{r}
dep %>%
na.omit() %>%
group_by(age_cat, culture) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T)) %>%
ggplot(aes(x = age_cat, y = avg_brood, fill = age_cat)) +
geom_col(position = "dodge", alpha = 0.5, color = "black") +
scale_fill_manual(values = my_palette, labels = c("Under 21", "21-25", "25-30", "30-35", "35-40", "40-45", "45-50", "50-55", "55-60", "Over 60")) +
scale_x_discrete(limits=c("teens", "earlytwens", "latetwens", "earlythirs", "latethirs", "earlyfors", "latefors", "earlyfifs", "latefifs", "sixties")) +
facet_grid(culture ~ ., scales = "free_y", space = "free", labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) +
scale_y_continuous(limits = c(0, 14), breaks = seq(0, 14, by = 2.5)) +
labs(title = "Brooding by Age Category and Culture",
x = NULL,
y = "Average Brooding Total",
fill = "Age Category") +
theme(plot.title = element_text(size = 12, face = "bold"),
plot.subtitle = element_text(size = 10),
axis.title = element_text(size = 10),
axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

```


```{r}
```


```{r}
dep %>% 
  na.omit() %>% 
  group_by(culture, depression_total) %>% 
  summarize(avg_brood = mean(brooding_total, na.rm = T)) %>% 
  ggplot(aes(x = depression_total, y = avg_brood, fill = factor(culture))) +
  geom_col(position = "dodge", alpha = 0.6) +
  scale_fill_manual(values = my_palette, labels = c("Euro-Australian", "Malaysian")) +
  labs(title = "Brooding by Depression Level and Culture",
       x = "Depression Level",
       y = "Brooding Total",
       fill = "Culture") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))


```


```{r}
dep %>% 
  na.omit() %>% 
  group_by(culture, anxiety_total) %>% 
  summarize(avg_brood = mean(brooding_total, na.rm = T)) %>% 
  ggplot(aes(x = anxiety_total, y = avg_brood, fill = factor(culture))) +
  geom_col(position = "dodge", alpha = 0.6) +
  scale_fill_manual(values = my_palette, labels = c("Euro-Australian", "Malaysian")) +
  labs(title = "Brooding by Anxiety Level and Culture",
       x = "Anxiety Level",
       y = "Brooding Total",
       fill = "Culture") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))


```

```{r}




```



```{r}

```

```{r}


```


```{r}
dep %>% 
  group_by(culture) %>% 
  na.omit() %>% 
  summarize(avg_brood = mean(brooding_total), avg_dep = mean(depression_total), avg_anx = mean(anxiety_total)) %>% 
  ggplot(aes(x = avg_dep , y = avg_brood, fill = culture))+
  geom_col(alpha=0.5)+
  scale_fill_manual(values=my_palette)+
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size=10)) +
  labs(title = "Brooding by Culture",
       x = "Culture",
       y = "Average Brooding",
       fill = "Culture")

```

```{r}
levels(dep$age_cat)

```


```{r}
glimpse(ptsd)
```




```{r}
glimpse(ptsd)
```

```{r}
names(ptsd)
```



