---
output: 
  html_document: 
    keep_md: yes
---

Based on 2022 paper, "https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0278328" Does culture moderate relationships between rumination and symptoms of depression and PTSD? 
Brooding/RRS rates & culture with: 

Akshanth:      religion, gender, education(1-5; change to factors) 

JKA:           depression, anxiety, age, trauma type for PTSD

-overall rumination rates
 -difference by age... to overal rumination rates (brooding total)
 
 Data codes for variables of interest from paper:

For both: First 109 participant ids are Euro-Australian, next 144 are Malaysian

For depression data:

2 Cultures:  0,1 = Euro-Australian, Malaysian (convert to factors)

Gender (convert to factors):              
 M = 1
 F = 2
Declined to state = 5  

Education (Highest level achieved):
Secondary = 1
Post-secondary = 2
Undergrad degree = 3
Graduate degree = 4
Other = 5

Religion codes:
None = 1
Christian = 2
Muslim = 3
Buddhist/Taoist = 4
Hindu = 5
Other (incl Sikh, Agnostic) = 6

PTSD dataset:

Culture = Groups 1,2  = Euro-Australian, Malaysian (change to factors)

Index trauma type (what the PTSD is thought to be in response to):

Accident/serious injury/illness = 1
Non-sexual assault/abuse = 2
Sexual assault/abuse = 3
Witnessing death = 4
War/natural disaster = 5
Other = 6       

HADS = Hospital Anxiety and Depression Scale: 

had_sdep = depression subscale total
had_sanx = anxiety subscale total
had_stotal = overall HADS total

pts_dtotal = overall score on PTSD assessment scale

RRS-B = Brooding scale - want sum of all RRS scores

>we want to look at pts (pcl 1 is q 1) total and if that value is above a certian value, then they have ptsd>> mutate boolean t/f 
>rrs is the brooding score sum the total and *maybe* append to depress to see total brooding score with depress  

So we are looking at rumination rates in the Y axis, and X axis is relig, gender, education...
and we have a facet of culture
```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)

```

```{r}
depress<-read_csv(here(
  #"R_project",
  "Rumination Depression.csv")) %>% clean_names
```


```{r}
post_disorder<-read_csv(here(
  #"R_project",
  "Rumination PTSD.csv")) %>% clean_names
```



```{r}
naniar::miss_var_summary(depress)
```


```{r}
glimpse(post_disorder)
```



Making the dataset readable

```{r}

#write.csv( d, "Depression_data", row.names=FALSE)

d<-depress %>% 
  select(-age,-brooding_total,-anxiety_total,-depression_total) %>% 
  mutate_if(is.numeric,as.character )
d<-d %>% mutate(age=depress$age)
d<-d %>% mutate(brooding_total=depress$brooding_total)
d<-d %>% mutate(anxiety_total=depress$anxiety_total)
d<-d %>% mutate(depression_total=depress$depression_total)
new_education<-d %>% select(education) %>% 
mutate_if( is.character, as.factor)

```
  To their names
```{r}

  
d["culture"][d["culture"]==0]<-"Australian"
d["culture"][d["culture"]==1]<-"Malaysian"  
d["gender"][d["gender"]==1]<-"M"
d["gender"][d["gender"]==2]<-"F"
d["gender"][d["gender"]==5]<-"Declined to State"
d["education"][d["education"]==1]<-"Secondary"
d["education"][d["education"]==2]<-"Post-secondary"
d["education"][d["education"]==3]<-"Undergrad degree"
d["education"][d["education"]==4]<-"Graduate degree"
d["education"][d["education"]==5]<-"Other"
d["religion_coded"][d["religion_coded"]==1]<-"None"
d["religion_coded"][d["religion_coded"]==2]<-"Christian"
d["religion_coded"][d["religion_coded"]==3]<-"Muslim"
d["religion_coded"][d["religion_coded"]==4]<-"Buddhist/Taoist"
d["religion_coded"][d["religion_coded"]==5]<-"Hindu"
d["religion_coded"][d["religion_coded"]==6]<-"Other (incl Sikh, Agnostic)"
# d<-d %>% 
#   select(-education) %>% mutate(education=new_education$education)

d <- d %>% 
  mutate(age_cat = case_when(age < 21 ~ "teens",
                             age >= 21 & age < 25 ~ "earlytwens",
                             age >= 25 & age < 30 ~ "latetwens",
                             age >= 30 & age < 35 ~ "earlythirs",
                             age >= 35 & age < 40 ~ "latethirs",
                             age >= 40 & age < 45 ~ "earlyfors",
                             age >= 45 & age < 50 ~ "latefors",
                             age >= 50 & age < 55 ~ "earlyfifs",
                             age >= 55 & age < 60 ~ "latefifs",
                             age >= 60 ~ "sixties"))
d
```

```{r}
names(d ) 
```

```{r}
d %>% 
  filter(!is.na(age)) %>% 
  ggplot(aes(x=brooding_total, y=age , fill=culture))+geom_dotplot(binaxis = "y",stackdir = "center", na.rm=T, binwidth = 1.000000000000001)+facet_wrap(culture~.)+coord_flip()+xlab("Total Brooding Score")+labs("Dotplot of Age versus its Brooding Score")
```

```{r}
d %>% 
  filter(!is.na(age)) %>% 
  ggplot(aes(x=age, y=brooding_total , color=culture))+geom_point(na.rm = T)+facet_wrap(culture~.)
```

```{r}
d %>% 
  filter(!is.na(age)) %>% 
  ggplot(aes(x=age, y=brooding_total , fill=culture))+geom_col(na.rm = T)+facet_wrap(culture~.)
```

```{r}
d %>% 
  filter(!is.na(age)) %>% 
 ggplot(aes(x=age, fill=culture))+geom_histogram(binwidth = 10, color="black")+facet_wrap(culture~.)+labs(title="Histogram of Age")
```


```{r}
d %>% 
  filter(!is.na(age)) %>% 
  ggplot(aes(x=gender, y= brooding_total, fill=culture))+geom_col(na.rm=T)+facet_wrap(culture~.)+labs(title="Gender vs Brooding Total")
```

```{r}
d %>% 
  filter(!is.na(religion_coded)) %>% 
  ggplot(aes(x=religion_coded, y= brooding_total, fill=education))+geom_col(na.rm=T, position = "dodge")+facet_wrap(culture~.)+coord_flip()+labs(title="Brooding Scores for Each of the Religions")
```


```{r}
d %>% 
  filter(!is.na(religion_coded)&religion_coded=="None") %>% 
    ggplot(aes(x=religion_coded, y= brooding_total, fill=education)) +geom_col(na.rm=T, position = "dodge") +facet_wrap(culture~.)+coord_flip()+labs(title="Brooding Score vs Religion: None")
```

```{r}
d %>% 
  filter(!is.na(religion_coded)&religion_coded=="Muslim") %>% 
    ggplot(aes(x=religion_coded, y= brooding_total, fill=education)) +geom_col(na.rm=T, position = "dodge") +facet_wrap(culture~.)+coord_flip()+labs(title="Brooding Score vs Religion: Muslim")
```

```{r}
d %>% 
  filter(!is.na(religion_coded)&religion_coded=="Hindu") %>% 
    ggplot(aes(x=religion_coded, y= brooding_total, fill=education)) +geom_col(na.rm=T, position = "dodge") +facet_wrap(culture~.)+coord_flip()+labs(title="Brooding Score vs Religion: Hindu")
```






```{r}
d %>% 
  filter(!is.na(culture)) %>% 
  ggplot(aes(x=depression_total, y=brooding_total, fill=culture))+geom_smooth(method="loess", na.rm = T, )+facet_wrap(culture~.)+labs(title="Sum of Depression Scores vs Brooding Scores")
      
```

```{r}
d %>% 
  filter(!is.na(culture)) %>% 
  ggplot(aes(x=anxiety_total, y=brooding_total, fill=culture))+geom_smooth( method = "loess", formula = "y~x")+facet_wrap(culture~.)+labs(title="Sum of Anxiety Scores vs Brooding Scores")
```
>below here is the work of JK abrams

```{r}
library(here)
library(tidyverse)
library(naniar)
library(janitor)
library(skimr)
library(ggplot2)
library(ggthemes)
library(RColorBrewer)
library(paletteer)
library(tidyr)
```

```{r}
my_palette <- paletteer::paletteer_d(("ggprism::colors"))
```

```{r}
dep<-read_csv("jka_r_project/rum_depress_jka.csv") %>% 
            clean_names 
```

```{r}
names(dep)
```


```{r}
ptsd<- read_csv("jka_r_project/rum_ptsd_jka.csv") %>% 
  clean_names 
```

```{r}
names(ptsd)
```
Tidying ptsd data

```{r}
glimpse(ptsd)
```

```{r}
ptsd_long <- ptsd %>%
  pivot_longer(cols = -c(participantnumber, timecode, had_stotal, age, gender, education, ethnicity, relgion, religioncode, group, traumacoding, pts_dtotal, had_sdep, had_sanx),
    names_to = "scale",
    values_to = "score",
    names_pattern = "(.*)"
  )
```


```{r}
ptsd_long <- ptsd_long %>%
  separate(scale, into = c("test", "q"), sep = "(?<=\\D)(?=\\d)", convert = TRUE)
ptsd_long
```
Preparing brooding/age/culture data for depression dataset

```{r}
dep$culture <- as.factor(dep$culture)
class(dep$culture)
```

Making sure the "culture" code is actually 0 for Euro-Australian and 1 for Malaysian - 
checking average ages against table in the paper

```{r}
dep %>% 
  filter(culture==0) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```

```{r}
dep %>% 
  filter(culture==1) %>% 
  summarize(max_age = max(age, na.rm =T),
            min_age = min(age,na.rm =T),
            avg_age = mean(age,na.rm =T),
            total = n())
```

Making age categories
```{r}
dep <- dep %>% 
  mutate(age_cat = case_when(age < 21 ~ "18 - 21",
                             age >= 21 & age < 25 ~ "21 - 25",
                             age >= 25 & age < 30 ~ "25 - 30",
                             age >= 30 & age < 35 ~ "30 - 35",
                             age >= 35 & age < 40 ~ "35 - 40",
                             age >= 40 & age < 45 ~ "40 - 45",
                             age >= 45 & age < 50 ~ "45 - 50",
                             age >= 50 & age < 55 ~ "50 - 55",
                             age >= 55 & age < 60 ~ "55 - 60",
                             age >= 60 ~ "Over 60"))

```


```{r}
class(dep$age_cat)
```

Checking the data before plotting
```{r}
dpage <- dep %>%
  group_by(culture, age_cat) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep") %>%
  select(culture, age_cat, avg_brood)
  
dpage$age_cat <- factor(dpage$age_cat, levels=c(
    "18 - 21",
    "21 - 25",
    "25 - 30",
    "30 - 35",
    "35 - 40",
    "40 - 45",
    "45 - 50",
    "50 - 55",
    "55 - 60",
    "Over 60"))

dpage
```

Reordering column order by age, 
and make them the same scale (0-14)
before plotting

```{r}
dpage %>% 
  filter(!is.na(age_cat)) %>% 
  ggplot(aes(x = age_cat, y = avg_brood, fill = age_cat)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) +
  scale_x_discrete(limits=c(
    "18 - 21",
    "21 - 25",
    "25 - 30",
    "30 - 35",
    "35 - 40",
    "40 - 45",
    "45 - 50",
    "50 - 55",
    "55 - 60",
    "Over 60"))+
  ylim(0,14)+
  labs(title = "Brooding by Age Category and Culture",
      x = NULL,
      y = "Average Brooding Score",
      fill = "Age Category") +
  theme(plot.title = element_text(size = 12, face = "bold"),
  plot.subtitle = element_text(size = 10),
  axis.title = element_text(size = 10),
  axis.text.x = element_text(angle = 45, hjust = 1))
```
Preparing depression/brooding/culture data from depression dataset

```{r}
dp2 <- dep %>%
group_by(culture, depression_total) %>%
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep") %>%
  select(culture, depression_total, avg_brood)
  
dp2$depression_total <- factor(dp2$depression_total)
```

```{r}
dp2 %>% 
  filter(!is.na(depression_total)) %>% 
  ggplot(aes(x = depression_total, y = avg_brood, fill = depression_total)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian"))) 
```
Fixing labels
```{r}
dp2 %>% 
  filter(!is.na(depression_total)) %>% 
  ggplot(aes(x = depression_total, y = avg_brood, fill = depression_total)) +
  geom_col(alpha = 0.5, color = "black")+
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian")))+ 
  ylim(0,18)+
labs(title = "Brooding by Depression Level and Culture",
       x = "Total Depression Score",
       y = "Average Brooding Score",
       fill = "Total Depression Score") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))
```
Preparing brooding/anxiety/culture data from depression dataset

```{r}
danx <- dep %>% 
  group_by(culture, anxiety_total) %>% 
  summarize(avg_brood = mean(brooding_total, na.rm = T), .groups ="keep")

danx$anxiety_total <- factor(danx$anxiety_total)
```

```{r}
danx <- danx %>% 
  filter(!is.na(anxiety_total) & 
         !is.na(avg_brood) & 
         !is.na(culture))

danx <- danx[-35,]

ggplot(danx, aes(x = anxiety_total, y = avg_brood, fill = anxiety_total)) +
  geom_col(alpha = 0.5, color = "black") +
  facet_grid(culture ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(culture = c(`0` = "Euro-Australian", `1` = "Malaysian")))+ 
  scale_y_continuous(limits = c(0, 20))+
   labs(title = "Brooding by Anxiety Level and Culture",
  x = "Total Anxiety Score",
  y = "Average Brooding Score",
  fill = "Total Anxiety Score") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8))
```
Preparing brooding/depression/culture data from ptsd dataset: 

brooding_total = total sum of rrs (repetitive rumination scores) for each participant
had_sdep = total depression score for each (Hospital Anxiety & Depression Scale)
had_sanx = total anxiety score for each
pts_dtotal= total score on PTSD scale  
traumacoding = code for initial trauma leading to PTSD
group = culture: 1 for Euro-Australian, 2 for Malaysian

```{r}
glimpse(ptsd_long)
```

Computing the sums of rrs scores (total brooding) for each participant, ptsd dataset 


Comibine culture(group) and participant number into a single "id"
```{r}
ptsd_long$id <- paste(ptsd_long$group, ptsd_long$participantnumber, sep = "_")
```

```{r}
glimpse(ptsd_long)
```


Computing sum of rrs scores
```{r}
ptsd_rrs_sum <- ptsd_long %>%
  filter(test == "rrs") %>%
 filter(!is.na(score)) %>%
 group_by(id) %>%
 summarise(rrs_sum = sum(score))
ptsd_rrs_sum
```

Separating the id back into participant & culture for plotting

```{r}
ptsd_sum_sep <- ptsd_rrs_sum %>% 
  separate (id, into = c("group", "participantnumber"), sep = "_", convert = TRUE) 
ptsd_sum_sep  
```

```{r}
ptsd_sum_sep$group <- as.factor(ptsd_sum_sep$group)
ptsd_sum_sep$participantnumber <- as.factor(ptsd_sum_sep$participantnumber)
ptsd_sum_sep$rrs_sum <- as.factor(ptsd_sum_sep$rrs_sum)
ptsd_long$participantnumber <- as.factor(ptsd_long$participantnumber)
```

```{r}
glimpse(ptsd_sum_sep)
```
Adding "had_sdep" to ptsd_sum_sep

```{r}
# Get unique values of group and participantnumber
unique_groups <- unique(ptsd_sum_sep$group)
unique_participants <- unique(ptsd_sum_sep$participantnumber)
```


```{r}
# remove duplicates from ptsd_sum_sep
ptsd_sum_sep <- distinct(ptsd_sum_sep, participantnumber, .keep_all = TRUE)
```

```{r}
# remove duplicates from ptsd_long
ptsd_long <- distinct(ptsd_long, participantnumber, had_sdep, .keep_all = TRUE)
glimpse(ptsd_long)
```

```{r}
glimpse(ptsd_sum_sep)
```



```{r}
# join the datasets
ptsd_sum_sep <- left_join(ptsd_long, ptsd_sum_sep, by = "participantnumber")
```


```{r}
str(ptsd_sum_sep$group.x)
```

```{r}
ptsd_sum_sep$group.x <- factor(ptsd_sum_sep$group.x)
class(ptsd_sum_sep$group.x)
```
Removing extra columns for ptsd_sum_sep after the join

```{r}
#ptsd_sum_sep <- subset(ptsd_sum_sep, select = -c(group.x.x, group.y))

```

```{r}
ptsd_sum_sep <- ptsd_sum_sep %>% 
  rename(group = group.x) %>% 
  slice(-220)
```


Plotting the scatterplot

```{r}
# Plot the scatter plot
ggplot(ptsd_sum_sep, aes(x = as.numeric(had_sdep), y = as.numeric(rrs_sum), color = group)) +
  geom_point(size = 0.75) +
  scale_color_manual(values = my_palette) +
  facet_grid(group ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(group = c(`1` = "Euro-Australian", `2` = "Malaysian")))+ 
  ylim(0,40)+
  labs(title = "Brooding by Depression Score and Culture, PTSD",
  x = "Total Depression Score",
  y = "Total Brooding Score",
  fill = "Culture") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    legend.position = "none")

```

```{r}

```

Converting trauma numbers to abbreviated codes for initial trauma:

AI = accident/injury/serious illness
NSA = non-sexual abuse
SA = sexual abuse
CD = witnessing death
WND = war/natural disaster
OTH = other

```{r}
ptsd_t1 <- ptsd_long %>% 
  filter(between(as.numeric(as.character(traumacoding)), 1, 6)) %>% 
  mutate(index_t = case_when (traumacoding == 1 ~ "AI",
                              traumacoding == 2 ~ "NSA",
                              traumacoding == 3 ~ "SA",
                              traumacoding == 4 ~ "CD",
                              traumacoding == 5 ~ "WND",
                              traumacoding == 6 ~ "OTH"))
```



```{r}
ptsd_sum_sep$participantnumber <- as.factor(ptsd_sum_sep$participantnumber)
class(ptsd_sum_sep$participantnumber)
```


Adding traumacoding to ptsd_sum_sep

```{r}
ptsd_sum_sep <- distinct(ptsd_sum_sep, participantnumber, .keep_all = TRUE)
ptsd_t1 <- distinct(ptsd_t1, participantnumber, .keep_all = TRUE)
```


```{r}
ptsd_sum_sep <- left_join(ptsd_sum_sep, ptsd_t1 %>% 
                             select(participantnumber, index_t), by = "participantnumber")
```


Get counts for each trauma type to check with the plot

```{r}
ptsd_sum_sep %>% 
  count(index_t, group) 
```

```{r}
ptsd_sum_sep %>% 
  group_by(index_t) %>% 
  count(index_t,rrs_sum)
```



```{r}
ptsd_sum_sep %>% 
  filter(!is.na(rrs_sum) & !is.na(index_t))%>%
  ggplot(aes(x = index_t, y = as.numeric(rrs_sum),  color = index_t)) +
  geom_point(na.rm = TRUE)+
  facet_grid(group ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(group = c(`1` = "Euro-Australian", `2` = "Malaysian"))) +
  ylim(0,40)+
  labs(title = "Brooding by Initial Trauma and Culture, PTSD",
      x = "Initial Trauma",
      y = "Brooding Score",
      color = "Initial Trauma",
      fill = "Initial Trauma") +
  theme(plot.title = element_text(size = 12, face = "bold"),
  plot.subtitle = element_text(size = 10),
  axis.title = element_text(size = 10),
  axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(limits = c("AI", "CD", "NSA", "SA", "WND", "OTH")) +
  scale_color_discrete(limits = c("AI", "CD", "NSA", "SA", "WND", "OTH"),
                       labels = c("Accident/Injury/Illness", "Witnessing Death", "Non-Sexual Abuse", "Sexual Abuse", "War/Natural Disaster", "Other"))

```
There were 10 participant numbers with no index trauma information, 5 Euro-Australian and 5 Malaysian. 



Preparing data for plotting brooding/PTSD level/culture

Making PTSD level categories

```{r}
ptsd_sum_sep <- ptsd_sum_sep %>% 
  mutate(ptsdlev = case_when(pts_dtotal <= 10 ~ "Unnoticeable",
                             pts_dtotal >= 11 & pts_dtotal < 20 ~ "Very Mild",
                             pts_dtotal >= 21 & pts_dtotal < 30 ~ "Mild",
                             pts_dtotal >= 31 & pts_dtotal < 40 ~ "Significant",
                             pts_dtotal >= 41 & pts_dtotal < 50 ~ "Marked",
                             pts_dtotal >= 51 & pts_dtotal < 60 ~ "Very Marked",
                             pts_dtotal >= 60 ~ "Severe"))
```


```{r}
ptsd_sum_sep %>% 
  group_by(ptsdlev) %>% 
  count(ptsdlev,rrs_sum)
```


```{r}
ptsd_sum_sep$ptsdlev <- factor(ptsd_sum_sep$ptsdlev, levels = c("Unnoticeable", "Very Mild", "Mild", "Moderate", "Significant", "Marked", "Severe"))
ggplot(ptsd_sum_sep[!is.na(ptsd_sum_sep$ptsdlev),], aes(x = ptsdlev, y = as.numeric(rrs_sum), color = factor(ptsdlev))) +
  geom_point(size = 1) +
  scale_color_manual(values = my_palette) +
 scale_color_discrete(breaks = c(
  "Unnoticeable",
  "Very Mild",
  "Mild",
  "Moderate",
  "Significant",
  "Marked",
  "Severe"),
    labels = c(
  "Unnoticeable",
  "Very Mild",
  "Mild",
  "Moderate",
  "Significant",
  "Marked",
  "Severe"
))+
   facet_grid(group ~ ., scales = "free_y", space = "free", 
             labeller = labeller(group = c(`1` = "Euro-Australian", `2` = "Malaysian"))) + 
   ylim(0, 45) +
  labs(title = "Brooding by PTSD Level and Culture, PTSD",
       x = "PTSD Level",
       y = "Total Brooding Score",
       color = "PTSD Level",
       fill = "PTSD Level") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))


```



```{r}
 ggplot(ptsd_sum_sep, aes(x = as.numeric(had_sanx), y = as.numeric(rrs_sum), color = factor(group))) +
  geom_point(size = 0.75) +
  scale_color_manual(values = my_palette) +
  facet_grid(group ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(group = c(`1` = "Euro-Australian", `2` = "Malaysian")))+ 
  ylim(0,40)+
  labs(title = "Brooding by Anxiety Score and Culture, PTSD",
  x = "Total Anxiety Score",
  y = "Total Brooding Score",
  fill = "Culture") + 
  theme(plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    legend.position = "none")
```
Preparing data for brooding/age category/culture, PTSD


Making age categories
```{r}
ptsd_sum_sep <- ptsd_sum_sep %>% 
  mutate(age_cat = case_when(age < 21 ~ "18 - 21",
                             age >= 21 & age < 25 ~ "21 - 25",
                             age >= 25 & age < 30 ~ "25 - 30",
                             age >= 30 & age < 35 ~ "30 - 35",
                             age >= 35 & age < 40 ~ "35 - 40",
                             age >= 40 & age < 45 ~ "40 - 45",
                             age >= 45 & age < 50 ~ "45 - 50",
                             age >= 50 & age < 55 ~ "50 - 55",
                             age >= 55 & age < 60 ~ "55 - 60",
                             age >= 60 ~ "Over 60"))
```

```{r}
glimpse(ptsd_sum_sep)
```

```{r}
ptsd_sum_sep$age_cat <- factor(ptsd_sum_sep$age_cat, levels=c(
    "18 - 21",
    "21 - 25",
    "25 - 30",
    "30 - 35",
    "35 - 40",
    "40 - 45",
    "45 - 50",
    "50 - 55",
    "55 - 60",
    "Over 60"))
```

```{r}
str(ptsd_sum_sep)

```

Reordering column order by age, 
and make them the same scale (0-14)
before plotting

```{r}
as.factor(ptsd_sum_sep$age)
```


```{r}
ptsd_sum_sep %>% 
  filter(!is.na(age_cat) & !is.na(age))%>%
  ggplot(aes(x = age, y = as.numeric(rrs_sum),  color = age_cat)) +
  geom_point(na.rm = TRUE)+
  facet_grid(group ~ .,
             scales = "free_y", 
             space = "free", 
             labeller = labeller(group = c(`1` = "Euro-Australian", `2` = "Malaysian"))) +
  #scale_x_discrete(limits=c(
    #"18 - 21",
    #"21 - 25",
    #"25 - 30",
    #"30 - 35",
    #"35 - 40",
    #"40 - 45",
    #"45 - 50",
    #"50 - 55",
    #"55 - 60",
    #"Over 60"))+
  ylim(0,14)+
  labs(title = "Brooding by Age Category and Culture, PTSD",
      x = "Age Category",
      y = "Brooding Score",
      color = "Age Category",
      fill = "Age Category") +
  theme(plot.title = element_text(size = 12, face = "bold"),
  plot.subtitle = element_text(size = 10),
  axis.title = element_text(size = 10),
  axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
glimpse(ptsd_sum_sep)
```

```{r}
naniar::where_na(ptsd_sum_sep$age)
```


