---
title: "R Cheat Sheet"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    theme: spacelab
    number_sections: yes
date: "2023-01-28"
---


# Refrence text for formatting and publishing website in R
[For publishing in R](https://bookdown.org/yihui/rmarkdown/)

# Locating where you are and other basic operations
Getting working directory using the function `getwd()`
```{r}
getwd()
```
## Doing aritmatic

You can do arithmatic in R. Order of operations do apply.
Basic Operations: +, -, /, *
Note (**) is basically the exponent (^) function

## Creating a vector
you can create a vector by `x<- c(observation1, obs2, obs3,... )`
this uses the concatonate function which is c()
```{r}
x <- c(4, 6, 8, 5, 6, 7, 7, 7)
```
## Installing packages
You can install things called packages that would help you utilize other functions of R
Use the function `library (package_name)`
```{r}
library("tidyverse")
```
# Data types, Checking Data Types
There are five frequently used `classes` of data: 1. numeric, 2. integer, 3. character, 4. logical, 5. complex.
```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```
You can check if these using the `is.{classType}()` function
You can also change class types by the `as.{classType}( {variableName} )` function

In R, there can also be data values that don't exist, called NA's

# Statiscal knowledge

