---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("tidyverse")
library("purrr")
library("here")
library("jpeg")
library("janitor")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.
```{r}

malaria<-read_csv(here("lab9", "data","malaria.csv")) %>% clean_names()
malaria
```

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.
```{r}
malaria_2.0<-malaria %>% 
pivot_longer(-c(data_date, location_name,submitted_date,province,district,newid), names_to = "ranges", values_to = "rdt")

#you could also do pivot_longer(cols=star_with(asdhaskldask)<-this would have all the coloumns that are moved are the ones that start with it)
malaria_2.0

#select represents the data in the order that you put the data in
#like 
```

3. Make the data tidy and store them as a new object.
```{r}
malaria_2.0

```

4. Which district had the highest *total* number of cases on July 30, 2020?
```{r}
malaria_2.0 %>% 
  group_by(district) %>% 
  filter(data_date=="2020-07-30",ranges=="malaria_tot") %>% 
  
  summarise(rdt_total=sum(rdt, na.rm = T)) %>% #FINDING THE SUM IN A GROUP IS DONE BY USING THE SUM FUNCTION!!!!
  arrange(-rdt_total)
```

