---
title: "Lab 4 practice MIDTERM AAAAAA"
output: html_document
date: "2023-01-19"
---

1. In a new R Markdown document, load the tidyverse and a package called `palmerpenguins`.
2. What are the dimensions of the dataset `penguins`?
3. What are the names of the variables in the `penguins` dataset?
4. How many individuals were sampled on each island in the `penguins` dataset?
5. What is the mean body mass for all individuals in the `penguins` dataset?
1)

*note we use tidyverse because it makes things easier to read and understand
tools->check packages updates=> make sure your packages are up to date 

```{r}


library("tidyverse")
library("palmerpenguins")
```

2)
```{r}
dim(penguins)
```

3)

```{r}
colnames(penguins)
```
4)
TABLEfunction (table()) is what we use for counts for qualitative data (like not numbers)


```{r}
table(penguins$island)
```
5)
```{r}
bdwght<-filter(penguins, !is.na(body_mass_g))

mean(bdwght$body_mass_g)
view(penguins$body_mass_g)

mean(penguins$body_mass_g, na.rm = T)
```



