---
title: "Lab 6 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
library("tidyverse")
library("janitor")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What is a pipe? Why are they useful in R?

Pipe is a way to put/apply funcitons in R. They allow us to work inside a certain workflow, as they apply under a certain funciton    
condense coding into one block and saves 


this allows us change coding quickly

2. What is the shortcut for making a pipe?

shift+ cmd+ m

cntrl+shift+m for PC

3. What is the difference between `select`, `filter`, and `mutate`?

Select allows you to choose a certain coloumn, filter allows you to change a certain row, and mutate allows you to do a lot: add rows, change rows, change observations, append coloumsn.

4. What are the logistics of midterm 1?

50 mins Solo, turn in for 5 point, then 50 mins for group work

### Warm-up
1. Open the data `ecs21351-sup-0003-SupplementS1.csv`
```{r}
supplemental<-readr::read_csv("data/ecs21351-sup-0003-SupplementS1.csv", skip=2)
```

2. Clean the names of the variables
```{r}
clean_names(supplemental)
tolower(colnames(supplemental))

```

3. Explore the data using a function of your choice
```{r}
str(supplemental)
```

4. Are there more veggie, omni, insecivore, or nectar feeders in the data?
```{r}
tabyl(supplemental$Diet)%>% 
    arrange(-n)
```

5. Which families are Veggie?
```{r}
supplemental %>% 
  filter(diet=="Vegetation")
  select(family, diet)
```

